Efficiently Handling Guards in the Low Level Design of RPython's tracing JIT

### 2012-07-31 talk/vmil2012/paper.tex
typo
Add a figure about the control flow in case of patched and unpatched guard failures and refer to the figure in the text

\bivab{example for low-level resume data goes here}

@@

As in previous sections the underlying idea for the design of guards is to have
a fast on-trace profile and a potentially slow one in the bail-out case where
the execution takes one of the side exits due to a guard failure. At the same
time the data stored in the backend needed to rebuild the state should be be
as compact as possible to reduce the memory overhead produced by the large
number of guards\bivab{back this}.

@@

Once the bridge has been compiled the trampoline method stub is redirected to
the code of the bridge. In future if the guard fails again it jumps to the code
compiled for the bridge instead of bailing out. Once the guard has been
compiled and attached to the loop the guard becomes just a point where
control-flow can split. The loop after the guard and the bridge are just
conditional paths. \todo{add figure of trace with trampoline and patched guard to a bridge}
%* Low level handling of guards
%   * Fast guard checks v/s memory usage
%   * memory efficient encoding of low level resume data
%   * fast checks for guard conditions
%   * slow bail out
%
% section Guards in the Backend (end)

@@

In~\cite{Gal:2006} Gal et. al describe that in the HotpathVM they experimented
with having one generic compensation code block, like the RPython JIT, that
uses a register variable mapping to restore the interpreter state. Later this
was replaced by generting compensation code for each guard which produced a
lower overhead in their benchmarks. HotpathVM also records secondary traces
starting from failing guards that are connected directly to the original trace.
Secondary traces are compiled by first restoring the register allocator state to
the state at the side exit. The information is retrieved from a mapping stored
in the guard that maps machine level registers and stack to Java level stack
and variables.

>>>>

\todo{example for low-level resume data showing how the current encoding works?}

@@

As in previous sections the underlying idea for the design of guards is to have
a fast on-trace profile and a potentially slow one in the bail-out case where
the execution takes one of the side exits due to a guard failure. At the same
time the data stored in the backend needed to rebuild the state needs to be as
compact as possible to reduce the memory overhead produced by the large number
of guards, the numbers in Figure~\ref{fig:backend_data} illustrate that the
compressed encoding currently has about 25\% of the size of of the generated
instructions on x86.

@@

Once the bridge has been compiled the guard that led to compiling the birdge is
patched to redirect control flow to the bridge in case the check fails. In
future if the guard fails again it jumps to the code compiled for the bridge
instead of bailing out. Once the guard has been compiled and attached to the
loop the guard becomes just a point where control-flow can split. The loop
after the guard and the bridge are just conditional paths.
Figure~\ref{fig:trampoline} shows a digram of a compiled loop with two guards,
Guard \#1 jumps to the trampoline, loads the \texttt{low level resume data} and
then calls the compensation code, whereas Guard \#2 has already been patched
and directly jumps to the corresponding bridge. The bridge also contains two
guards that work based on the same principles.
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/loop_bridge.pdf}
\caption{Trace control flow in case of guard failures with and without bridges}
\label{fig:trampoline}
\end{figure}
%* Low level handling of guards
%   * Fast guard checks v/s memory usage
%   * memory efficient encoding of low level resume data
%   * fast checks for guard conditions
%   * slow bail out
%
% section Guards in the Backend (end)

@@

In~\cite{Gal:2006} Gal et. al describe that in the HotpathVM they experimented
with having one generic compensation code block, like the RPython JIT, that
uses a register variable mapping to restore the interpreter state. Later this
was replaced by generating compensation code for each guard which produced a
lower overhead in their benchmarks. HotpathVM also records secondary traces
starting from failing guards that are connected directly to the original trace.
Secondary traces are compiled by first restoring the register allocator state to
the state at the side exit. The information is retrieved from a mapping stored
in the guard that maps machine level registers and stack to Java level stack
and variables.
